\chapter{Project Management in Information Systems}

\section{Introduction}
Project management in Information Systems (IS) involves planning, executing, and monitoring projects to achieve organizational goals. A project is a planned sequence of activities with a defined start and end, differing from ongoing business processes.

\section{Why Projects are Undertaken}
Organizations initiate projects for two primary reasons:
\begin{itemize}
    \item To take advantage of a business opportunity.
    \item To solve a business problem.
\end{itemize}

\section{Phases of Project Management}
Project management consists of four key phases:

\subsection{Project Initiation}
\begin{itemize}
    \item Assess project size, scope, and complexity.
    \item Define objectives and key stakeholders.
\end{itemize}

\subsection{Project Planning}
\begin{itemize}
    \item Define clear, discrete activities required to complete the project.
    \item Establish a \textbf{Work Breakdown Structure (WBS)}: Dividing the project into manageable tasks.
    \item Estimate task duration using three estimates:
        \begin{itemize}
            \item Optimistic Time (To).
            \item Most Likely Time (Tm).
            \item Pessimistic Time (Tp).
        \end{itemize}
\end{itemize}

\subsection{Project Execution}
\begin{itemize}
    \item Implement the baseline plan.
    \item Monitor progress and manage changes.
    \item Revise completion dates as necessary.
\end{itemize}

\subsection{Project Closure}
\begin{itemize}
    \item Bring the project to completion.
    \item Assess success or failure.
    \item Address \textbf{project starvation}, where budget cuts prevent project completion.
\end{itemize}

\section{Project Control and Monitoring}
\textbf{Project control} ensures alignment with goals through:
\begin{itemize}
    \item Monitoring progress.
    \item Addressing schedule slippage.
    \item Managing scope creep (unplanned expansion of project commitments).
\end{itemize}

\subsection{Brooks’ Law}
\textbf{Brooks' Law} states that \textbf{"adding manpower to a late software project makes it later"} due to increased communication overhead.

\subsection{Parkinson’s Law}
\textbf{Parkinson’s Law} states that \textbf{"work expands to fill the time available for completion"}.

\section{Project Scheduling Techniques}
\subsection{Gantt Charts}
\begin{itemize}
    \item Developed by Henry Gantt in 1917.
    \item Provides a visual representation of project tasks and progress.
\end{itemize}

\subsection{Network Techniques: PERT and CPM}
\begin{itemize}
    \item \textbf{Program Evaluation and Review Technique (PERT)}: Structures projects as networks of activities.
    \item \textbf{Critical Path Method (CPM)}: Identifies the longest sequence of tasks that determines project duration.
\end{itemize}

\subsection{Critical Path Analysis}
\begin{itemize}
    \item Identifies tasks that directly impact project completion time.
    \item \textbf{Slack time}: The amount of time a task can be delayed without affecting the overall project timeline.
\end{itemize}

\section{AI in Project Management}
Artificial Intelligence (AI) enhances project management by:
\begin{itemize}
    \item Predicting project risks using historical data.
    \item Automating task allocation and scheduling.
    \item Improving decision-making with AI-driven insights.
\end{itemize}

\section{Case Study: AI-Powered Project Scheduling}
A global IT firm implemented AI-based scheduling tools to manage software development projects. The system analyzed past project data, optimized task assignments, and reduced delays by 25\%.

\section{Multiple Choice Questions (MCQs)}
\begin{enumerate}
    \item What defines a project?
    \begin{enumerate}
        \item[A.] Ongoing activities without a defined end
        \item[B.] A planned sequence of activities with a defined objective
        \item[C.] Daily routine business operations
        \item[D.] A one-time activity with no objective
    \end{enumerate}
    \textbf{Answer: B}

    \item What is the first phase of project management?
    \begin{enumerate}
        \item[A.] Execution
        \item[B.] Planning
        \item[C.] Initiation
        \item[D.] Closure
    \end{enumerate}
    \textbf{Answer: C}

    \item What is the purpose of a Work Breakdown Structure (WBS)?
    \begin{enumerate}
        \item[A.] To merge multiple projects
        \item[B.] To divide the project into smaller, manageable tasks
        \item[C.] To determine software requirements
        \item[D.] To eliminate project risks
    \end{enumerate}
    \textbf{Answer: B}

    \item Which of the following best describes scope creep?
    \begin{enumerate}
        \item[A.] Project goals shrinking over time
        \item[B.] Uncontrolled expansion of project requirements
        \item[C.] A fixed project schedule
        \item[D.] Increasing project costs intentionally
    \end{enumerate}
    \textbf{Answer: B}

    \item What does Brooks' Law state?
    \begin{enumerate}
        \item[A.] Adding manpower to a late software project makes it later
        \item[B.] Every project has a critical path
        \item[C.] Increasing budget speeds up project completion
        \item[D.] All projects must be completed in phases
    \end{enumerate}
    \textbf{Answer: A}

    \item Which of the following is NOT a project management phase?
    \begin{enumerate}
        \item[A.] Initiation
        \item[B.] Execution
        \item[C.] File storage
        \item[D.] Closure
    \end{enumerate}
    \textbf{Answer: C}

    \item What is the purpose of PERT analysis?
    \begin{enumerate}
        \item[A.] To determine the longest sequence of project tasks
        \item[B.] To define organizational hierarchy
        \item[C.] To manage financial transactions
        \item[D.] To eliminate project risks
    \end{enumerate}
    \textbf{Answer: A}

    \item What is slack time in project management?
    \begin{enumerate}
        \item[A.] The amount of time a task can be delayed without affecting the project
        \item[B.] The maximum time required for a project
        \item[C.] The amount of time wasted on unnecessary meetings
        \item[D.] The extra cost of hiring new employees
    \end{enumerate}
    \textbf{Answer: A}

    \item How does AI contribute to project management?
    \begin{enumerate}
        \item[A.] Predicting project risks
        \item[B.] Automating scheduling
        \item[C.] Improving decision-making
        \item[D.] All of the above
    \end{enumerate}
    \textbf{Answer: D}

    \item What is a major benefit of using Gantt charts?
    \begin{enumerate}
        \item[A.] They make projects more expensive
        \item[B.] They help visualize project timelines and progress
        \item[C.] They eliminate the need for team meetings
        \item[D.] They replace all project management software
    \end{enumerate}
    \textbf{Answer: B}
\end{enumerate}
